Internal Gravity Wave Breaking in White Dwarf Binaries

### 2019-08-26 writeups/Su_IGW_break.tex
small edits

Slices of
$\overline{U}, \hat{F}$ across $z$ at various times $t$ are shown in
\autoref{fig:nl_fluxes}. While the behavior of $\overline{U}$ seems to conform
qualitatively with the predictions of \autoref{ss:crit_layer}, the behavior of
$\hat{F}$ exhibits two salient features: (i) the incident flux seems to
fluctuate greatly with time, and (ii) there seems to be a small transmitted
feature at many of the later times. We will discuss further these features in
\autoref{ss:reflectivity}, after first analyzing the propagation of the critical
layer.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{plots/nl_fluxes.png}
    \caption{Plot of $\overline{U}(z, t)$ and $\hat{F}(z, t)$ respectively at
    various times over $z$ in our $\mathrm{Re} = 1024$ simulation.
    $\overline{U}, \hat{F}$ follow their definitions in \autoref{eq:mean_flow}
    and \autoref{eq:sbm_def}. $\overline{U}$ is given in units of
    $\overline{U}_c = \frac{\omega}{k_{x}}$ the critical horizontal flow
    velocity (see \autoref{ss:crit_layer}). The propagation of the critical
    layer towards lower $z$ and sharp deposition of $F$ at the critical layer
    are evident.}\label{fig:nl_fluxes}
\end{figure}

@@

Since fluid instabilities are local, $\mathrm{Ri}$ must be measured in the
immediate vicinity of a fluid parcel. Thus, we first assign an $\mathrm{Ri}$ for
every $x$ in the critical layer, then take the median as $\mathrm{Ri}$ for the
entire layer. To avoid noisiness, the local $\mathrm{Ri}$ is computed using the
vertical distance over which the local $u_x$ increases from $0.3 \times$ its
critical value to its critical value. The value $0.3$ effectively excludes the
self-acceleration induced by the IGW\@. This can be written
\begin{align}
    z_{CL, \min}(x, t) &= \argmin_\zeta
        \z{z: u_x(x, \zeta, t) > 0.3\overline{U}_c},\nonumber\\
    z_{CL, \max}(x, t) &= \argmax
        \zeta \z{z: u_x(x, \zeta, t) < \overline{U}_c},\nonumber\\
    \mathrm{Ri}(t) &\equiv
        \med_x\p{\frac{N^2 \p{z_{CL, \max} - z_{CL, \min}}^2}{(0.7
            \overline{U}_c)^2}}.\label{ss:ri_med_def}
\end{align}
To understand the variation in $\mathrm{Ri}$ over $x$, we can also plot using
the minimum over $x$ (the maximum is significantly noisier). Both of these are
shown in \autoref{fig:nl_f_ri}. The quick evolution of $\mathrm{Ri}$ to its
saturated value is reflective of the fact that IGW are \emph{anti-diffusive}.
This property is a simple consequence of the IGW dispersion relation
\autoref{eq:disp_rel}, which is approximately $\omega k_z \approx Nk_x$ such
that as an IGW propagates into a shear flow, $\omega$ decreases and $k_z$
increases, enhancing dissipation and further driving mean flow acceleration.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{plots/nl_f_ri.png}
    \caption{Local Richardson number of the flow at the critical layer over time
    as defined in \autoref{ss:ri_med_def}. Dotted red lines demarcate the
    minimum and median of $\mathrm{Ri}$ over $x$. These numbers effectively
    measure the mean and spread in width of the critical layer over $x$. Note
    that $\mathrm{Ri} \sim \frac{1}{4}$ corresponds to the KHI, so this plot
    suggests the shear at the critical layer does not steepen past the KHI
    onset.}\label{fig:nl_f_ri}
\end{figure}

>>>>

Slices of $\overline{U}, \hat{F}$ across $z$ at the times depicted in
\autoref{fig:snapshots} are shown in \autoref{fig:nl_fluxes}. While the behavior
of $\overline{U}$ seems to conform qualitatively with the predictions of
\autoref{ss:crit_layer}, the behavior of $\hat{F}$ exhibits two notable
features: (i) the incident flux seems to fluctuate greatly with time, and (ii)
there seems to be a small transmitted feature at many of the later times. We
will discuss further these features in \autoref{ss:reflectivity}, after first
analyzing the propagation of the critical layer.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{plots/nl_fluxes.png}
    \caption{Plot of $\overline{U}(z, t)$ and $\hat{F}(z, t)$ respectively at
    the same times in \autoref{fig:snapshots} over $z$ in our $\mathrm{Re} =
    1024$ simulation. $\overline{U}, \hat{F}$ follow their definitions in
    \autoref{eq:mean_flow} and \autoref{eq:sbm_def}. $\overline{U}$ is given in
    units of $\overline{U}_c = \frac{\omega}{k_{x}}$ the critical horizontal
    flow velocity (see \autoref{ss:crit_layer}). The propagation of the critical
    layer towards lower $z$ and sharp deposition of $F$ at the critical layer
    are evident.}\label{fig:nl_fluxes}
\end{figure}

@@

Since fluid instabilities are local, $\mathrm{Ri}$ must be measured in the
immediate vicinity of a fluid parcel. Thus, we first assign an $\mathrm{Ri}$ for
every $x$ in the critical layer, then take the median as $\mathrm{Ri}$ for the
entire layer. To avoid noisiness, the local $\mathrm{Ri}$ is computed using the
vertical distance over which the local $u_x$ increases from $0.3 \times$ its
critical value to its critical value. The value $0.3$ effectively excludes the
self-acceleration induced by the IGW\@. This can be written
\begin{align}
    z_{CL, \min}(x, t) &= \argmin_\zeta
        \z{z: u_x(x, \zeta, t) > 0.3\overline{U}_c},\nonumber\\
    z_{CL, \max}(x, t) &= \argmax
        \zeta \z{z: u_x(x, \zeta, t) < \overline{U}_c},\nonumber\\
    \mathrm{Ri}(t) &\equiv
        \med_x\p{\frac{N^2 \p{z_{CL, \max} - z_{CL, \min}}^2}{(0.7
            \overline{U}_c)^2}}.\label{ss:ri_med_def}
\end{align}
To understand the variation in $\mathrm{Ri}$ over $x$, we can also plot using
the minimum over $x$ (the maximum is significantly noisier). Both of these are
shown in \autoref{fig:nl_f_ri}. The quick evolution of $\mathrm{Ri}$ to its
saturated value is reflective of the fact that IGW are \emph{anti-diffusive}.
This property is a simple consequence of the IGW dispersion relation
\autoref{eq:disp_rel}, which is approximately $\omega k_z \approx Nk_x$ such
that as an IGW propagates into a shear flow, $\omega$ decreases and $k_z$
increases, enhancing dissipation and further driving mean flow acceleration.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{plots/nl_f_ri.png}
    \caption{Local Richardson number of the flow at the critical layer over time
    as defined in \autoref{ss:ri_med_def}. The dotted red and solid green lines
    denote respectively the minimum and median of $\mathrm{Ri}$ over $x$ (the
    maximum greatly exceeds the vertical scale of this plot). These numbers
    effectively measure the mean and spread in width of the critical layer over
    $x$. Note that $\mathrm{Ri} \sim \frac{1}{4}$ corresponds to the KHI, so
    this plot suggests the shear at the critical layer does not steepen past the
    KHI onset.}\label{fig:nl_f_ri}
\end{figure}
